Determinism, Complexity, and Predictability in Computer Performance

### 2014-03-18 model.tex
finished the model.tex rewrite

In this section, we describe the four different forecasting schema
used in this study, as well as the error metric used to evaluate their
predictive accuracy.  These methods include:
\begin{itemize}
\item The \emph{random-walk} method, which uses the previous value in
  the observed signal as the forecast,

@@

\item The \emph{LMA} method, which uses a near-neighbor forecast
  strategy on a dynamical reconstruction of the signal.
\end{itemize}
ARIMA is based on standard autoregressive techniques.  LMA is designed
to capture and exploit the deterministic structure of a signal from a
nonlinear dynamical system.  The first two methods, somewhat
surprisingly, often outperform these more-sophisticated prediction
strategies in the case of highly complex signals, as discussed below.

@@

A random-walk predictor simply uses the last observed measurement as
the forecast: that is, the value $px_i$ at time $i$ is predicted using
the relation $$px_i = x_{i-1}$$ The prediction strategy that we refer
to using the term ``\naive'' averages the prior observations to
generate the forecast: $$px_i = \sum_{j=1}^{i-1}\frac{x_j}{i-1}$$
While both of these methods are simplistic, they are not without
merit.  For a time series near the high end of the complexity
spectrum---i.e., one that possesses very little predictive
structure---these two methods can actually be the best choice.  In
forecasting currency exchange rates, for instance, sophisticated
econometrics-based prediction models fail to consistently outperform
the random-walk method~\cite{rwMeese,rwCCE}.  These signals are
volatile, noisy, and possess very little predictive structure, but
they do not---on the average---vary a great deal, so the random-walk
method's strategy of simply guessing the last known value is not a bad
choice.  If a signal has a unimodal distribution with low variance,
the \naive ~prediction strategy will perform quite well---even if the
signal is highly complex---simply because the mean is a good
approximation of the future behavior.  Moreover, a temporal average
effects a low-pass filtering operation.  In a signal with very little
predictive structure, then, the \naive ~prediction strategy
effectively mitigates the complexity.

Both of these methods have significant weaknesses, however.  Because
they do not model the temporal patterns in the data, or even the
distribution of its values, they cannot track changes in that
structure.  This causes them to fail in a number of important
situations.  Random-walk strategies are a particularly bad choice for
time series that change drastically at every time step.  In the worst
case---a large-amplitude square wave whose period is equivalent to
twice the sample time---a random-walk prediction would be exactly 180
degrees out of phase with the true continuation.  The \naive ~method
would be a better choice in this situation, since it would always
predict the mean.  It would, however, perform poorly when a signal has
a number of long-lived regimes that have significantly different
means.  In this situation, the inertia of the \naive ~method's
accumulating mean is a liability and the agility of the random-walk
method is an advantage, since it can respond quickly to regime shifts.

Of course, methods that could capture and exploit the geometry of the
data, or its temporal patterns, would be far more effective in the
situations described in the previous paragraph.  The ARIMA and LMA
methods introduced in Sections~\ref{sec:lma} and~\ref{sec:arima} are
designed to do exactly that.  Note, though, that periodic patterns
appear only in signals that are at the low end of the complexity
spectrum.  If a signal contains little predictive structure, forecast
strategies like ARIMA and LMA have nothing to work with and thus will
often be outperformed by the two simple strategies described in this
section.  This is explored further in Sections~\ref{sec:accuracy}
and~\ref{sec:results}.

@@

A simple and yet powerful way to capture and exploit the structure of
data is to fit a hyperplane to the dataset and then use that it to
make predictions.  The roots of this approach date back to Yule's 1927
invention of the autoregressive schema~\cite{weigend93}, which
forecasts the next time step through a weighted average of past
observations: $$px_i = \sum a_i x_i$$ The weighting coefficients $a_i$
are computed using a polynomial of degree $p$ that is fit to the past
$p+1$ values of $x_i$.  To account for noise in the data, one can add
a so-called ``moving average'' term to the model; to remove
nonstationarities, one can detrend the data using a differencing
operation.  A strategy that incorporates all three of these features
is called a \emph{nonseasonal ARIMA model}.  If evidence of periodic
structure is present in the data, a \emph{seasonal ARIMA model}, which
adds a sampling operation that filters out periodicities, can be a
good choice.

There is a vast amount of theory and literature regarding the
construction and use of models of this type; we refer the reader to
\cite{davislinearts} for an in-depth exploration.  For the purposes of
this paper, where the goal is to explore the relationship between
predictability and complexity across a broad array of forecast
strategies, seasonal ARIMA models are a good exemplar of the class of
linear forecasting strategies.  Fitting such a model to a dataset
involves choosing values for the various free parameters in the
autoregressive, detrending, moving average, and filtering
terms---e.g., the $p$ in the equation at the beginning of this
section.  We employ the automated fitting techniques described
in~\cite{autoARIMA} to accomplish this.  This procedure uses
sophisticated methods---KPSS unit-root tests~\cite{KPSSunit}, a
customization of the Canova-Hansen test~\cite{Canova1995}, and the
Akaike information criterion~\cite{akaike1974}, conditioned on the
maximized likelihood of the model fitted to the detrended data---to
select good values for the free parameters of the ARIMA model.

ARIMA forecasting is a common and time-tested procedure.  Its
adjustments for seasonality, nonstationarity, and noise make it an
appropriate choice for short-term predictions of time-series data
generated by a wide range of processes.  If information is being
generated and/or transmitted in a nonlinear way, however, a global
linear fit is inappropriate and ARIMA forecasts will be inaccurate.
Another weakness of this method is prediction horizon, since an ARIMA
forecast is guaranteed to converge to the mean after some number of
predictions, depending on model order.  To avoid this issue, we build
forecasts in a stepwise fashion: i.e., fit the model to the existing
data, use it to perform a one-step prediction, rebuild the model using
the latest observations, and iterate until the desired prediction
horizon is reached.  (For consistency, we take the same approach with
the other three models in this study as well, even though this amounts
to artificially hobbling LMA.)

@@

Delay-coordinate embedding allows one to reconstruct a dynamical
system's full state-space dynamics from a scalar time-series
measurement---provided that some conditions hold regarding that data.
Specifically, if the underlying dynamics and the measurement
function---the mapping from the unknown state vector $\vec{X}$ to the
observed value $x_i$---are both smooth and generic,
Takens~\cite{takens} formally proves that the delay-coordinate map
\[
F(\tau,m)(\vec{X}) = ([x_{i} ~ x_{i+\tau} ~ \dots ~x_{i+m\tau}])
\]
from a $d$-dimensional smooth compact manifold $M$ to
$\mathbb{R}^{2d+1}$ is a diffeomorphism on $M$---in other words, that
the reconstructed dynamics and the true (hidden) dynamics have the
same topology.  This is an extremely powerful result: among other
things, it means that one can model the full system dynamics without
measuring, or even knowing, every one of its state variables.

@@

Geometric structure in these kinds of plots is an indication of
structure in the underlying information generation/transmission
process.  The dynamical systems community has developed a number of
methods that leverage this structure to generate predictions
(e.g.,~\cite{weigend-book,casdagli-eubank92,Smith199250}).  One of the
most straightforward of these is the ``Lorenz method of analogues''
(LMA), which is essentially nearest-neighbor prediction in the
embedded\footnote{Lorenz's original formulation used the full system
  state space; the method was later extended to embedded dynamics by
  Kantz~\cite{kantz-lma-citation}.} space~\cite{lorenz-analogues}.

Even this simple algorithm---which builds predictions by finding the
nearest neighbor in the embedded space of the given point, then taking
that neighbor's path as the forecast---provides accurate forecasts
when the generating process is a deterministic dynamical system.
Since LMA does not rest on an assumption of linearity, as ARIMA does,
it can be applied to data produced by both linear and nonlinear
processes.  If the underlying generating process is not deterministic,
however, it can perform poorly.  In Figure~\ref{fig:embedding}(b), for
instance, little structure is visible, so LMA may not work well.  More
structure appears to be present in~Figure~\ref{fig:embedding}(c), but
this reconstruction also appears to contain some noise.

The question as to how much structure is present in a
reconstruction---and how much of that structure can be captured and
used by LMA---is apropos of the central question treated in this
paper.  It may be that \gcc has some redundancy that LMA cannot
exploit, or that the structure in \svdfive is effectively obfuscated,
from the standpoint of the LMA method, by noise.  By quantifying the
balance between normalized redundancy (that is, predictive structure)
and entropy for these real-valued time series, we can begin to answer
these questions.

@@

To study the relationship between predictability and complexity, we
use these four methods to generate predictions of all 360 traces
described in Section~\ref{sec:methods}, then calculate the error
between the predictions and the true continuations.  Specifically, we
split each time series into two pieces: the first 90\%, referred to as
the ``initial training" signal and denoted $\{x_i\}_{i=1}^{n}$, and
the last 10\% known as the ``test" signal $\{c_j\}_{j=n+1}^{k+n+1}$.
The initial training signal is used to build the model, following the
procedures described in the previous section; that model is used to
generate a prediction of the value of $x_{n+1}$, which is then
compared to the true continuation, $c_{n+1}$.  The model is then
rebuilt using $\{x_i\}_{i=1}^{n+1}$ and the process repeats $k$ times,
out to the end of the observed time series.  This ``one step
prediction'' process is not technically necessary in the LMA method,
which can generate arbitrary-length predictions, but the performance
of the other three methods used here will degrade if the associated
models are not periodically rebuilt.  In order to make the comparison
fair, we used an iterative one-step prediction schema {\sl for all
  four methods}.  This has the slightly confusing effect of causing
the ``test'' signal to be used both to assess the accuracy of each
model and for periodic refitting.

To assess prediction accurary, we calculate the mean absolute scaled
error (MASE) between the true and predicted signals:
% 
$$MASE = \sum_{j=n+1}^{k+n+1}\frac{|c_j-p_j| }{\frac{k}{n-1}\sum^n_{i=2}|x_{i}-x_{i-1}|}$$
% 
This error method was introduced in \cite{MASE} as a ``generally
applicable measurement of forecast accuracy without the problems seen
in the other measurements."  Note that the scaling term in the
denominator
% 
% of the MASE value
% $$\frac{1}{n-1}\sum^n_{i=2}|X_{i,obs}-X_{i-1,obs}|$$ 
% 
is the average in-sample forecast error for a random-walk prediction
$px_i=x_{i-1}$ over a signal of length $k$.  That is, MASE is a
normalized measure.  A MASE value below one means that the prediction
error was, on the average, smaller than the error of a one-step
in-sample random-walk forecast on the same data.  Analogously, a MASE
value greater than one means that the corresponding prediction method
did \emph{worse}, on average, than a one-step in-sample random-walk
method.  We chose this error metric because it allows for fair
comparison across varying methods, prediction horizons, and signal
scales.

Table \ref{tab:error} provides the MASE distributions {\color{red}[[Joshua: Ryan, Is this the right word? we give mean $\pm$ std. dev but some have very skewed right tails]]} for each of the eight signals and three prediction strategies, these are averaged over 15 runs of each signal-method combination. For later comparison, Table \ref{tab:error} also has the complexity measure we introduce in Section  \ref{sec:meaComplex}.

To aid in a conceptual understanding of this error metric we provide example forecasts in Figure~\ref{fig:forecast-example}. In plots like this we plot $(c_j,p_j)$, visually points close to the line $y=x$, i.e. the line $p_j = c_j$ are more accurate. Forecasts which stay close to the line are more accurate than forecasts that stray away from it.

\begin{figure}[htbp]
  \centering
      \begin{subfigure}{0.33\columnwidth}
    \includegraphics[width=\columnwidth]{figs/colMeanForecast.png}
    \caption{\col\\ na\"ive }
    \label{fig:gccMEAN}
  \end{subfigure}%
   \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\columnwidth]{figs/gccMeanForecast.png}
    \caption{\gcc\\ na\"ive }
    \label{fig:gccMEAN}
  \end{subfigure}% 
     \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\columnwidth]{figs/svdfiveMeanForecast.png}
    \caption{\svdfive\\ na\"ive }
    \label{fig:gccMEAN}
  \end{subfigure}%      
  \\
    \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\columnwidth]{figs/colARIMAForecast.png}

    \caption{\col ARIMA}
    \label{fig:colARIMA}
  \end{subfigure}
  \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\columnwidth]{figs/gccARIMAForecast.png}
    \caption{\gcc \\ ARIMA }
    \label{fig:gccARIMA}
  \end{subfigure}%
  \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\columnwidth]{figs/svdfiveARIMAForecast.png}
    \caption{\svdfive \\ ARIMA }
    \label{fig:gccARIMA}
  \end{subfigure}%    
  \\

      \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\columnwidth]{figs/colLMAForecast.png}
    \caption{\col LMA}
    \label{fig:colLMA}
  \end{subfigure}
      \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\columnwidth]{figs/gccLMAForecast.png}
    \caption{\gcc LMA}
    \label{fig:gccLMA}
  \end{subfigure}  
        \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\columnwidth]{figs/svdfiveLMAForecast.png}
    \caption{\svdfive LMA}
    \label{fig:gccLMA}
  \end{subfigure}
  %\begin{subfigure}{0.5\textwidth}
  %  \includegraphics[width=1.0\textwidth]{figs/LMA_vs_ARIMA}
   \caption{
{\color{red}[[Joshua: Liz, I added \svdfive as it has been a persistent example up to this figure. I am a little worried about the 9x9 being too small in 1 column. We may need to make this figure span two columns, which I think is ok.]]}For each of these, we plot the predicted value $p_j$ against the correct value $c_j$. On this type of plot a perfect prediction lies exactly on the diagonal, that is the line $p_j = c_j$, e.g., (g) is a near perfect prediction where-as (e) is a poor prediction.}\label{fig:forecast-example}  
  %\end{subfigure}
\end{figure}

It may be helpful for the reader at this point to compare some of the MASE scores in Table \ref{tab:error} with the example forecasts in Figure~\ref{fig:forecast-example}. For example, if we compare Figures~ \ref{fig:forecast-example} (d) and (g), \col is almost perfectly predicted by LMA (close to the line $y=x$ and ARIMA is a poor predictor as the points are spread out around $y=x$. Comparing to the MASE values in Table~\ref{tab:error} we see that \col with LMA had a MASE of $0.05 \pm 0.002$, whereas \col with ARIMA had a MASE of $0.599 \pm 0.211$. So intuitively the better visual prediction received a much better MASE score (the smaller the better). One way of interpreting these are that LMA on \col is on average 20 times more effective than random walk and ARIMA on \col is only 2 times more effective than random walk.

The part that becomes tricky and which must be kept in mind is that this error metric is in comparison to random walk. If we compare visually for example the LMA prediction of \col and \svdfive (Figures~\ref{fig:forecast-example} (g) \& (i)) visually they look very similar in terms of accuracy; however, \col was forecast 20 times better than random walk where as \svdfive was forecast only 1.5 times better. The reason for this lies in the strengths/weaknesses of Random Walk discussed in Section~\ref{sec:simple}. The large oscillations in \col cause Random Walk to do poorly, MASE then gives a model that accounts for this oscillation (e.g., LMA) a boost, whereas \svdfive is a great signal for Random Walk so there is less improvement to be made, and less boost is given for example for correctly reacting to regime shift. A more in-depth discussion of this is presented in Section~\ref{sec:results}.

@@

% took out for space

>>>>

In this section, we describe the four different forecasting methods
used in this study, as well as the error metric used to evaluate their
predictive accuracy.  These methods include:
\begin{itemize}
\item The \emph{random-walk} method, which uses the previous value in
  the observed signal as the forecast,

@@

\item The \emph{LMA} method, which uses a near-neighbor forecast
  strategy on a dynamical reconstruction of the signal.
\end{itemize}
ARIMA is based on standard autoregressive techniques.  LMA is designed
to capture and exploit the deterministic structure of a signal from a
nonlinear dynamical system.  The \naive and random-walk methods,
somewhat surprisingly, often outperform these more-sophisticated
prediction strategies in the case of highly complex signals, as
discussed below.

@@

A random-walk predictor simply uses the last observed measurement as
the forecast: that is, the value $px_i$ at time $i$ is calculated
using the relation $$px_i = x_{i-1}$$ The prediction strategy that we
refer to using the term ``\naive'' averages the prior observations to
generate the forecast: $$px_i = \sum_{j=1}^{i-1}\frac{x_j}{i-1}$$
While both of these methods are simplistic, they are not without
merit.  For a time series near the high end of the complexity
spectrum---i.e., one that possesses very little predictive
structure---these two methods can actually be the best choice.  In
forecasting currency exchange rates, for instance, sophisticated
econometrics-based prediction models fail to consistently outperform
the random-walk method~\cite{rwMeese,rwCCE}.  These signals are
volatile, noisy, and possess very little predictive structure, but
they do not---on the average---vary a great deal, so the random-walk
method's strategy of simply guessing the last known value is not a bad
choice.  If a signal has a unimodal distribution with low variance,
the \naive ~prediction strategy will perform quite well---even if the
signal is highly complex---simply because the mean is a good
approximation of the future behavior.  Moreover, the \naive
~prediction strategy's temporal average effects a low-pass filtering
operation, which effectively mitigates the complexity in signals with
very little predictive structure.

Both of these methods have significant weaknesses, however.  Because
they do not model the temporal patterns in the data, or even the
distribution of its values, they cannot track changes in that
structure.  This causes them to fail in a number of important
situations.  Random-walk strategies are a particularly bad choice for
time series that change significantly at every time step.  In the
worst case---a large-amplitude square wave whose period is equivalent
to twice the sample time---a random-walk prediction would be exactly
180 degrees out of phase with the true continuation.  The \naive
~method would be a better choice in this situation, since it would
always predict the mean.  It would, however, perform poorly when a
signal has a number of long-lived regimes that have significantly
different means.  In this situation, the inertia of the \naive
~method's accumulating mean is a liability and the agility of the
random-walk method is an advantage, since it can respond quickly to
regime shifts.

Of course, methods that could capture and exploit the geometry of the
data, or its temporal patterns, would be far more effective in the
situations described in the previous paragraph.  The ARIMA and LMA
methods introduced in Sections~\ref{sec:arima} and~\ref{sec:lma} are
designed to do exactly that.  Note, though, that periodic patterns
appear only in signals that are at the low end of the complexity
spectrum.  If a signal contains little predictive structure, forecast
strategies like ARIMA and LMA have nothing to work with and thus will
often be outperformed by the two simple strategies described in this
section.  This is explored further in Sections~\ref{sec:accuracy}
and~\ref{sec:results}.

@@

A simple and yet powerful way to capture and exploit the structure of
data is to fit a hyperplane to the dataset and then use that it to
make predictions.  The roots of this approach date back to Yule's 1927
invention of the autoregressive schema~\cite{weigend93}, which
forecasts the next time step through a weighted average of past
observations: $$px_i = \sum_{j=1}^{i-1} a_j x_j$$ The weighting
coefficients $a_i$ are computed using a polynomial of degree $p$ that
is fit to the past $p+1$ values of $x_i$.  To account for noise in the
data, one can add a so-called ``moving average'' term to the model; to
remove nonstationarities, one can detrend the data using a
differencing operation.  A strategy that incorporates all three of
these features is called a \emph{nonseasonal ARIMA model}.  If
evidence of periodic structure is present in the data, a
\emph{seasonal ARIMA model}, which adds a sampling operation that
filters out periodicities, can be a good choice.

There is a vast amount of theory and literature regarding the
construction and use of models of this type; we refer the reader to
\cite{davislinearts} for an in-depth exploration.  For the purposes of
this paper, where the goal is to explore the relationship between
predictability and complexity across a broad array of forecast
strategies, seasonal ARIMA models are a good exemplar of the class of
linear predictors.  Fitting such a model to a dataset involves
choosing values for the various free parameters in the autoregressive,
detrending, moving average, and filtering terms---e.g., the $p$ in the
equation at the beginning of this section.  We employ the automated
fitting techniques described in~\cite{autoARIMA} to accomplish this.
This procedure uses sophisticated methods---KPSS unit-root
tests~\cite{KPSSunit}, a customization of the Canova-Hansen
test~\cite{Canova1995}, and the Akaike information
criterion~\cite{akaike1974}, conditioned on the maximized likelihood
of the model fitted to the detrended data---to select good values for
the free parameters of the ARIMA model.

ARIMA forecasting is a common and time-tested procedure.  Its
adjustments for seasonality, nonstationarity, and noise make it an
appropriate choice for short-term predictions of time-series data
generated by a wide range of processes.  If information is being
generated and/or transmitted in a nonlinear way, however, a global
linear fit is inappropriate and ARIMA forecasts will be inaccurate.
Another weakness of this method is prediction horizon, since an ARIMA
forecast is guaranteed to converge to the mean after some number of
predictions, depending on model order.  To sidestep this issue, we
build forecasts in a stepwise fashion: i.e., fit the model to the
existing data, use it to perform a one-step prediction, rebuild the
model using the latest observations, and iterate until the desired
prediction horizon is reached.  (For consistency, we take the same
approach with the other three models in this study as well, even
though this amounts to artificially hobbling LMA.)

@@

Delay-coordinate embedding~\cite{packard80,Sauer:1991lr,Takens:1981uq}
allows one to reconstruct a dynamical system's full state-space
dynamics from a scalar time-series measurement---provided that some
conditions hold regarding that data.  Specifically, if the underlying
dynamics and the measurement function---the mapping from the unknown
state vector $\vec{X}$ to the observed value $x_i$---are both smooth
and generic, Takens~\cite{Takens:1981uq} formally proves that the
delay-coordinate map
\[
F(\tau,m)(\vec{X}) = ([x_{i} ~ x_{i+\tau} ~ \dots ~x_{i+m\tau}])
\]
from a $d$-dimensional smooth compact manifold $M$ to
$\mathbb{R}^{2d+1}$ is a diffeomorphism on $M$---in other words, that
the reconstructed dynamics and the true (hidden) dynamics have the
same topology.  This is an extremely powerful result: among other
things, it means that one can model the full system dynamics without
measuring, or even knowing, every one of its state variables.

@@

Geometric structure in these kinds of plots is an indication of
structure in the information generation/transmission process the
produced the time series.  The dynamical systems community has
developed a number of methods that leverage this structure to generate
predictions (e.g.,~\cite{weigend-book,casdagli-eubank92,Smith199250}).
One of the most straightforward of these is the ``Lorenz method of
analogues'' (LMA), which is essentially nearest-neighbor prediction in
the embedded\footnote{Lorenz's original formulation used the full
  system state space; the method was later extended to embedded
  dynamics by Kantz~\cite{kantz-lma-citation}.}
space~\cite{lorenz-analogues}.  Even this simple algorithm---which
builds predictions by finding the nearest neighbor in the embedded
space of the given point, then taking that neighbor's path as the
prediction---provides accurate forecasts when the generating process
is a deterministic dynamical system.

Since LMA does not rest on an assumption of linearity (as ARIMA does)
it can handle both linear and nonlinear processes.  If the underlying
generating process is nondeterministic, however, it can perform
poorly.  In Figure~\ref{fig:embedding}(b), for instance, little
structure is visible, so LMA may not work well.  More structure
appears to be present in~Figure~\ref{fig:embedding}(c), but this
reconstruction also appears to contain some noise.  The question as to
how much structure is present in a reconstruction---and how much of
that structure can be captured and used by LMA---is apropos of the
central question treated in this paper.  It may be that \gcc has some
redundancy that LMA cannot exploit, or that the structure in \svdfive
is effectively obfuscated, from the standpoint of the LMA method, by
noise.  By quantifying the balance between normalized redundancy (that
is, predictive structure) and entropy for these real-valued time
series, as shown in Section~\ref{sec:results}, we can begin to answer
these questions.

@@

To study the relationship between predictability and complexity, we
use the four methods outlined above to generate predictions of all 360
traces described in Section~\ref{sec:methods}, then calculate the
error between the predictions and the true continuations.
Specifically, we split each time series into two pieces: the first
90\%, referred to as the ``initial training" signal and denoted
$\{x_i\}_{i=1}^{n}$, and the last 10\%, known as the ``test" signal
$\{c_j\}_{j=n+1}^{k+n+1}$.  The initial training signal is used to
build the model, following the procedures described in the previous
section; that model is used to generate a prediction of the value of
$x_{n+1}$, which is then compared to the true continuation, $c_{n+1}$.
The model is then rebuilt using $\{x_i\}_{i=1}^{n+1}$ and the process
repeats $k$ times, out to the end of the observed time series.  This
``one step prediction'' process is not technically necessary in the
LMA method, which can generate arbitrary-length predictions, but the
performance of the other three methods used here will degrade if the
associated models are not periodically rebuilt.  In order to make the
comparison fair, we used an iterative one-step prediction schema {\sl
  for all four methods}.  This has the slightly confusing effect of
causing the ``test'' signal to be used both to assess the accuracy of
each model and for periodic refitting.

Figure~\ref{fig:forecast-example} shows example forecasts made using
the \naive, ARIMA, and LMA methods for the \col, \gcc, and \svdfive
time series.
\begin{figure*}[htbp]
  \centering
      \begin{subfigure}{0.6\columnwidth}
    \includegraphics[width=\columnwidth]{figs/colMeanForecast.png}
    \caption{\col\\ na\"ive }
    \label{fig:gccMEAN}
  \end{subfigure}%
   \begin{subfigure}{0.6\columnwidth}
    \includegraphics[width=\columnwidth]{figs/gccMeanForecast.png}
    \caption{\gcc\\ na\"ive }
    \label{fig:gccMEAN}
  \end{subfigure}% 
     \begin{subfigure}{0.6\columnwidth}
    \includegraphics[width=\columnwidth]{figs/svdfiveMeanForecast.png}
    \caption{\svdfive\\ na\"ive }
    \label{fig:gccMEAN}
  \end{subfigure}%      
  \\
    \begin{subfigure}{0.6\columnwidth}
    \includegraphics[width=\columnwidth]{figs/colARIMAForecast.png}

    \caption{\col \\ ARIMA}
    \label{fig:colARIMA}
  \end{subfigure}
  \begin{subfigure}{0.6\columnwidth}
    \includegraphics[width=\columnwidth]{figs/gccARIMAForecast.png}
    \caption{\gcc \\ ARIMA }
    \label{fig:gccARIMA}
  \end{subfigure}%
  \begin{subfigure}{0.6\columnwidth}
    \includegraphics[width=\columnwidth]{figs/svdfiveARIMAForecast.png}
    \caption{\svdfive \\ ARIMA }
    \label{fig:gccARIMA}
  \end{subfigure}%    
  \\

      \begin{subfigure}{0.6\columnwidth}
    \includegraphics[width=\columnwidth]{figs/colLMAForecast.png}
    \caption{\col \\ LMA}
    \label{fig:colLMA}
  \end{subfigure}
      \begin{subfigure}{0.6\columnwidth}
    \includegraphics[width=\columnwidth]{figs/gccLMAForecast.png}
    \caption{\gcc \\ LMA}
    \label{fig:gccLMA}
  \end{subfigure}  
        \begin{subfigure}{0.6\columnwidth}
    \includegraphics[width=\columnwidth]{figs/svdfiveLMAForecast.png}
    \caption{\svdfive \\ LMA}
    \label{fig:gccLMA}
  \end{subfigure}
  %\begin{subfigure}{0.5\textwidth}
  %  \includegraphics[width=1.0\textwidth]{figs/LMA_vs_ARIMA}
   \caption{Predicted ($px_j$) versus true values ($c_j$) for
     forecasts of \col, \gcc, and \svdfive.  On this type of plot, a
     perfect prediction lies exactly on the diagonal.
%
{\color{red} Joshua, what would you think of adding a row of
  random-walk plots at the top?  The letters in the subcaptions---(a)
  (b) etc.---feel clunky here.  Finally, we should eventually change
  the label of the vertical axes on these plots to be whatever symbol
  we choose for the prediction.}
%
}
\label{fig:forecast-example}
  %\end{subfigure}
\end{figure*} 
In these images, the vertical axis is the prediction $px_j$ and the
horizontal axis is the true continuation $c_j$.  On such a plot, a
perfect prediction would lie on the diagonal.  LMA, for instance,
generates a very accurate prediction of the \col data, while ARIMA
does not.  [[We should comment on what a noise ball means on such a
    plot, as well as what a horizontal line means.]]

As a numerical measure of prediction accurary, we calculate the mean
absolute scaled error ($MASE$) between the true and predicted signals:
% 
$$MASE = \sum_{j=n+1}^{k+n+1}\frac{|c_j-x_j|
}{\frac{k}{n-1}\sum^n_{i=2}|x_{i}-x_{i-1}|}$$
% 
This error metric was introduced in \cite{MASE} as a ``generally
applicable measurement of forecast accuracy without the problems seen
in the other measurements."  $MASE$ is a normalized measure; the
scaling term in the denominator
% 
% of the MASE value
% $$\frac{1}{n-1}\sum^n_{i=2}|X_{i,obs}-X_{i-1,obs}|$$ 
% 
is the average in-sample forecast error for a random-walk prediction
over a signal of length $k$.  That is, $MASE<1$ means that the
prediction error in question was, on the average, smaller than the
error of a random-walk forecast on the same data.  Analogously,
$MASE>1$ means that the corresponding prediction method did
\emph{worse}, on average, than the random-walk method.  We chose this
error metric because it allows for fair comparison across varying
methods, prediction horizons, and signal scales.

$MASE$ scores for all 360 experiments are tabulated in the three
middle columns in Table~\ref{tab:error}.
\begin{table*}
  \begin{center}
  \begin{tabular}{|c|c|c|c|c|}
  \hline Signal & MASE LMA & MASE ARIMA &MASE na\"{i}ve & Weighted Permutation \\ 
  &  &  & &  Entropy \\ \hline

 \col           & $ 0.050 \pm0.002  $ & $0.599  \pm 0.211 $ & $0.571\pm0.002$&  $0.513 \pm 0.003$ \\

\gcc           & $ 1.530\pm 0.021$ & $1.837 \pm0.016 $ & $0.951 \pm 0.001$ & $0.943 \pm 0.001$ \\

\svdone     & $ 0.827\pm 0.076$ & $ 0.714\pm 0.075 $ & $2.676\pm4.328$&  $0.957 \pm 0.016$ \\

 \svdtwo    & $1.279 \pm0.020 $ & $2.163 \pm0.027 $ &  $3.054\pm0.040$ &   $0.846 \pm0.004$ \\

 \svdthree     & $0.619 \pm0.021 $ & $0.713 \pm 0.010 $ & $31.386\pm 0.282$ &  $0.716 \pm 0.006$ \\

 \svdfour     & $ 0.779\pm0.036 $ & $0.979 \pm0.032 $ & $2.661\pm0.074$ & $0.825 \pm 0.008$ \\

 \svdfive     & $ 0.718\pm 0.048 $ & $2.370  \pm 0.051 $ & $20.870 \pm 0.192$&  $0.678 \pm 0.007$ \\

 \svdsix     & $ 0.739\pm 0.068 $ & $ 1.438\pm 0.0610$ & $2.197\pm0.083$&  $0.748 \pm 0.011$ \\
  \hline
  \end{tabular}
  \end{center}
 \label{default}
 \caption{MASE scores and weighted permutation entropies for all eight
   processes studied in this paper}
 \label{tab:error}
  \end{table*}%
Comparing these with the geometry of the plots in
Figure~\ref{fig:forecast-example}, one can see some obvious
correspondences.  The average LMA $MASE$ score for the \col signals
was $0.050$, for instance, while ARIMA scored much worse ($0.599$).
That is, LMA performed roughly 20 times better on \col signals than a
random-walk predictor, while ARIMA only outperformed random walk by a
factor of 1.7 [[is this right?  there was a ``2'' here, but 1/0.599 is
    closer to 1.7]].  This is in accordance with the visual appearance
of the corresponding images in Figure~\ref{fig:forecast-example}.
While the correspondence between $MASE$ score and the visual
appearance of a $px$ versus $c$ plot is clear cut in this case, that
is not always so.  The plots of LMA predictions of \col and \svdfive
both lie near the diagonal, for instance, but the corresponding $MASE$
scores were very different: $0.050$ for \col and $0.718$ for \svdfive
(resp., 20 and 1.4 times better than random-walk forecasts of the same
signals).  This is a result of the normalization in the $MASE$ score.
Recall from Section~\ref{sec:simple} that the random-walk method
performs especially poorly on signals that oscillate rapidly.  \col
fits this description, so the random-walk error on this signal---the
denominator of the three $MASE$ scores in the first row of the
Table---is pathologically high, which skews those scores down.
Random-walk prediction is very effective for the\svdfive signal, on
the other hand, so the denominator is small and the $MASE$ scores are
skewes upwards.  Normalization is a two-edged sword here.  It
facilitates comparisons across signals of different types and lengths,
but its particular pathologies must be taken into account, as
discussed at more length in Section~\ref{sec:results}.

% Table \ref{tab:error} provides the $MASE$ distributions
% {\color{red}[[Joshua: Ryan, Is this the right word? we give mean $\pm$
%       std. dev but some have very skewed right tails]]} for each of
% the eight signals and three prediction strategies, these are averaged
% over 15 runs of each signal-method combination. For later comparison,
% Table \ref{tab:error} also has the complexity measure we introduce in
% Section \ref{sec:meaComplex}.

@@
